## 空间自相关

## NYDistricts是shp数据
## NewEvictions是csv数据
## boro_cd是可以把shp和csv合并在一起的关键column
```{r}
library(sf)
points_sf_joined <- NYDistricts %>%
  st_join(NewEvictions) %>%
  add_count(boro_cd) %>%
  janitor::clean_names() %>%
  #calculate area
  mutate(area=st_area(.)) %>%
  #then density of the points per ward
  mutate(density=n/area) %>%
  #select density and some other variables 
  dplyr::select(density, boro_cd, n)
```

```{r}
points_sf_joined <- points_sf_joined %>%                    
  group_by(boro_cd) %>%         
  summarise(density = first(density),
          boro_cd = first(boro_cd),
          plaquecount= first(n))
tm_shape(points_sf_joined) +
    tm_polygons("density",
        style="jenks",
        palette="PuOr",
        midpoint=NA,
        popup.vars=c("boro_cd", "density"),
        title="Evictions Density")
```
So, from the map, it looks as though we might have some clustering of blue plaques in the Manhatten so let’s check this with Moran’s I and some other statistics.

Before being able to calculate Moran’s I and any similar statistics, we need to first define a  W_ij spatial weights matrix.
```{r}
library(spdep)
```
```{r}
#First calculate the centroids of all Wards in New York City
coordsW <- points_sf_joined%>%
  st_centroid()%>%
  st_geometry()
  
plot(coordsW,axes=TRUE)
```
```{r}
#create a neighbours list
LWard_nb <- points_sf_joined %>%
  poly2nb(., queen=T)
```

```{r}
summary(LWard_nb)
```
Average number of links is 4.422535

```{r}
#plot the neighbours
plot(LWard_nb, st_geometry(coordsW), col="red")
#add a map underneath
plot(points_sf_joined$geometry, add=T)
```

```{r}
#create a spatial weights matrix from these weights
#Lward.lw <- LWard_nb %>%
#  nb2mat(., style="B")
#sum(Lward.lw)
#summary(Lward.lw)
```

#### Spatial autocorrelation

##### Moran's I

Now we have defined our W_ij matrix, we can calculate the Moran’s I and other associated statistics. However, Moran’s I requires a spatial weight list type object as opposed to matrix.
```{r}
Lward.lw <- LWard_nb %>%
  nb2listw(., style="C")
```

Moran’s I test tells us whether we have clustered values (close to 1) or dispersed values (close to -1)

```{r}
I_LWard_Global_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  moran.test(., Lward.lw)
I_LWard_Global_Density
```

##### Geary’s C
This tells us whether similar values or dissimilar values are clustering
```{r}
C_LWard_Global_Density <- 
  points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  geary.test(., Lward.lw)
C_LWard_Global_Density
```


##### Getis Ord General G
This tells us whether high or low values are clustering. If G > Expected = High values clustering; if G < expected = low values clustering
```{r}
G_LWard_Global_Density <- 
  points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  globalG.test(., Lward.lw)
G_LWard_Global_Density
```

So the global statistics are indicating that we have spatial autocorrelation of Blue Plaques in London:

The Moran’s I statistic = 0.59 (remember 1 = clustered, 0 = no pattern, -1 = dispersed) which shows that we have some distinctive clustering

The Geary’s C statistic = 0.46 (remember Geary’s C falls between 0 and 2; 1 means no spatial autocorrelation, <1 - positive spatial autocorrelation or similar values clustering, >1 - negative spatial autocorreation or dissimilar values clustering) which shows that similar values are clustering

The General G statistic = G > expected, so high values are tending to cluster.


We can now also calculate local versions of the Moran’s I statistic (for each Ward) and a Getis Ord G*_i statistic to see where we have hot-spots…
```{r}
#use the localmoran function to generate I for each ward in the city
I_LWard_Local_count <- points_sf_joined %>%
  pull(plaquecount) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()
I_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localmoran(., Lward.lw)%>%
  as_tibble()
#what does the output (the localMoran object) look like?
slice_head(I_LWard_Local_Density, n=5)
```

```{r}
points_sf_joined <- points_sf_joined %>%
  mutate(plaque_count_I = as.numeric(I_LWard_Local_count$Ii))%>%
  mutate(plaque_count_Iz =as.numeric(I_LWard_Local_count$Z.Ii))%>%
  mutate(density_I =as.numeric(I_LWard_Local_Density$Ii))%>%
  mutate(density_Iz =as.numeric(I_LWard_Local_Density$Z.Ii))
```

##### Mapping outputs

We'll set the breaks manually based on the rule that data points >2.58 or <-2.58 standard deviations away from the mean are significant at the 99% level (<1% chance that autocorrelation not present); >1.96 - <2.58 or <-1.96 to >-2.58 standard deviations are significant at the 95% level (<5% change that autocorrelation not present).
```{r}
breaks1<-c(-1000,-2.58,-1.96,-1.65,1.65,1.96,2.58,1000)
```

```{r}
MoranColours<- rev(brewer.pal(8, "RdGy"))
```

```{r}
tm_shape(points_sf_joined) +
    tm_polygons("plaque_count_Iz",
        style="fixed",
        breaks=breaks1,
        palette=MoranColours,
        midpoint=NA,
        title="Local Moran's I, Blue Plaques in London")
```
This map shows some areas in the centre of London that have relatively high scores, indicating areas with lots of blue plaques neighbouring other areas with lots of blue plaques.

the Getis Ord G*_i statisic for hot and cold spots
```{r}
Gi_LWard_Local_Density <- points_sf_joined %>%
  pull(density) %>%
  as.vector()%>%
  localG(., Lward.lw)
head(Gi_LWard_Local_Density)
```

```{r}
points_sf_joined <- points_sf_joined %>%
  mutate(density_G = as.numeric(Gi_LWard_Local_Density))
```

```{r}
GIColours<- rev(brewer.pal(8, "RdBu"))
#now plot on an interactive map
tm_shape(points_sf_joined) +
    tm_polygons("density_G",
        style="fixed",
        breaks=breaks1,
        palette=GIColours,
        midpoint=NA,
        title="Gi*, Blue Plaques in London")